---
output:
  pdf_document: default
  html_document: default
---
Technical Skills Assessment
Author: Shiraz Bheda

```{r global_options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)

getwd()
setwd("C://Users//shira//Desktop//Technical Skills Assessment_2022")

#install.packages("tree")
#install.packages("naniar")
#install.packages("ppcor")
library("tree")
library("dplyr")
library("naniar")
library(tidyverse)
library(randomForest)
library(ppcor)
library(dplyr)
```

```{r}
###load input files into a dataframe
df <- read.csv("clinical.csv", stringsAsFactors = F)

###assign "ID" column as the row names vector, remove from dataframe
rownames(df) <- df$ID
df$ID <- NULL

###rename certain columns for clarity
df <- df %>% 
       rename("Tumor.Stage" = "T",
              "Metastasis.To.Lymph.Nodes" = "N",
              "Distant.Metastases" = "M")
```

Pre-processing checklist

#1. Check that field separators are splitting metadata in clinical.csv input file as intended
```{r}
###check dimensions of dataframe
dim(df)
head(df)
tail(df)
```
2. Exploratory data analysis: Check for any misspellings, blank spaces, formatting inconsistencies
```{r}
###Column 'Survival.Months' has 3 instances of floating-point numbers (9.5 months)
table(df$Survival.Months)

###Column 'Stage' has an incorrect value that needs to be changed (1B instead of IB)
table(df$Stage)
df$Stage[df$Stage == '1B'] <- 'IB'
table(df$Stage)

###Column 'Primary.Site' has incorrect spellings ('Righ Upper Lobe')
table(df$Primary.Site)
df$Primary.Site[df$Primary.Site == 'Righ Upper Lobe'] <- 'Right Upper Lobe'
table(df$Primary.Site)
```

3. Sanity-checks: Make sure that information in clinical.csv file is logically consistent
```{r}
###(Assuming that mutations & genes referred to are all in coding regions) all Values in 
###'Num.Mutations' should not be lower than corresponding values in 'Num.Mutated.Genes'
###column
check1 <- df[df$Num.Mutations < df$Num.Mutated.Genes, ]
nrow(check1) ###No selected rows indicates logic is ok between two columns


###Since the data comes from de-identified databases in order to fulfill PHI compliance, 
###there is a possibility that the same patient is accidentally duplicated within a 
###dataset. So we add a check here to ensure that is not the case.
df$key<-apply(df,1,paste,collapse="")
check2 <- table(table(df$key))
check2 #all concatenated values are unique 

###Load genomics.csv file, check to make sure that the number of genes per sample agrees
###with data in clinical.csv file
df2 <- read.csv("genomics.csv", stringsAsFactors = F)
df3 <- as.data.frame(table(df2$ID))
rownames(df3) <- df3$Var1
df3$Var1 <- NULL
check3 <- merge(df, df3, by.x = 0, by.y = 0)

check3$Num.Mutated.Genes - check3$Freq #data in clinical.csv & genomics.csv files match

```
Feature selection: Drop columns with greater than 20% missing data, check features for high correlations
```{r}
###drop columns with missing data greater than 20%
df_dropped_features_with_missing_data <- df[,!names(df) %in% 
      c("Grade", "Tumor.Stage", "Metastasis.To.Lymph.Nodes", "Distant.Metastases",
        "Tumor.Size")]

###checking pearson correlation between Num.Mutated.Genes & Num.Mutations
check4 <- df_dropped_features_with_missing_data[, c(9, 10)]
pcor(check4, method = "pearson") #91% correlation indicates redundancy


###drop 'Num.Genes' and get rid of key
df_dropped_features_with_missing_and_redundant_data <- 
df_dropped_features_with_missing_data[,!names(
  df_dropped_features_with_missing_data) %in% 
      c("Num.Mutated.Genes", "key")]

```

```{r}
###Selection of inputs for the RF model
inputs <- df_dropped_features_with_missing_and_redundant_data

inputs <- as.data.frame(inputs)
inputs$Outcome <- as.character(inputs$Outcome)
inputs$Outcome <- as.factor(inputs$Outcome)
inputs$Age <- as.numeric(inputs$Age)
inputs$Num.Primaries <- as.numeric(inputs$Num.Primaries)
inputs$Radiation <- as.numeric(inputs$Radiation)
inputs$Stage <- as.factor(inputs$Stage)
inputs$Primary.Site <- as.factor(inputs$Primary.Site)
inputs$Histology <- as.factor(inputs$Histology)
inputs$Num.Mutations <- as.factor(inputs$Num.Mutations)

#####USING RF MODEL#####

#use bagging to avoid losing data while training model
set.seed(12345)
bag.tree.DN = randomForest(Outcome~., data = inputs, mtry = 3, ntree = 501,
                           importance = TRUE, na.action=na.roughfix, keep.inbag = TRUE,
                           norm.votes = FALSE, replace = FALSE)


###generate a visualization of per-category outcome error rate, as well as the OOB error
###rate, and how they change as the number of decision trees are increased
plot(bag.tree.DN, legend=TRUE)
```

```{r}
###generate summary statistics of model
summary(bag.tree.DN)
```

```{r}
###generate confusion matrix to evaluate final error rates
bag.tree.DN$confusion
```

```{r}
###save individual number of decision tree votes as a separate dataframe
bag.tree.DN_df <- as.data.frame(bag.tree.DN$votes)
dim(bag.tree.DN_df)

bag.tree.DN_df
```

```{r}
###export to csv format
write.csv(bag.tree.DN_df,
          "votes.csv")
print(row.names(bag.tree.DN_df))

```
